\normalfont\spacedallcaps{The economic consequences of different attitudes of a policy maker: a combined epidemiological-econometric study}

### 2021-03-16 Manuscript/manuscript.tex
PUB

Among the very large number of compartmental model available in the literature \cite{vynnycky2010introduction} we considered,  as the simulator of the epidemic diffusion the simplest one: the Susceptible-Infectious-Recovered-Dead (SIRD).  Our choice is motivated by the fact that this relative simple model provides the gross features of an epidemic \cite{fernandez2020estimating,al2020forecasting} with a relative small number of parameters \footnote{One in principle can consider a SIRD model,in which the parameters that are time-dependent, as done by Ferrari et al. in Ref. \cite{ferrari2020modelling} for the description of Italian situation.  On the other side it is possible to increase the complexity of the model with other compartments as done in the following paper \cite{giordano2020modelling} by Giordano et al.  Note that in this last case the resolution of 9 differential equation is required (accompanied by the estimation of a large number of parameters)}). The SIRD model,first proposed by Kermack and McKendrick in 1927 \cite{kermack1927contribution},  is given by the following set of differential equations\cite{vynnycky2010introduction}:  

@@

where S is the number of people that are still susceptible,  I the number of people that are infected and R people that are recovered while D are people that are death.  N denotes the total population,  that for the timing of this paper,  it will be considered fixed \footnote{Othervise, if longer horizontal timing is considered,  it is necessary to consider also a source term for the births and a well term for the natural deaths.  For further details see \cite{vynnycky2010introduction}}. On the other side $\beta$,$\gamma$ and $\mu$ are the parameters that shape the probability by which one individual in the model moves from a compartment to another: in particular $\beta$ is the probability to be infected, $\gamma $ the probability to recover and $\mu$ the probability to die (basically the lethality defined  as the probability to die conditionally to be ill).  Usually epidemiologist are interested in the ratio:  

@@

For each sales category we run the regression reported in the Eq.  \ref{DID-eq} on de-seasoned data reported in Fig \ref{Selling_plots} and \ref{Selling_plots_II}.  The coefficients obtained are reported in Tab.  \ref{selling_DID_table_delta} (the estimation of the other parameters is given in Supporting Information D).  As one can point out from the plots,  it is true that there is a pre-trend in the data before the event,  however as proved numerically in the Supporting Info B,  the slope of this pre-trend is almost two order of magnitude lesser with respect to the slope in the lock-down T$_{1}$ and $T_{2}$.  As consequence this pre-trend,  compared to the lock-down effect,  can be considered negligible.   In order to make a further control,  a placebo test was performed by choosing timing before the pandemic of COVID19.  As illustrated in the Supporting Info C,  this test was successful.  On the other side,  concerning $T_{3}$  (low lock-down),  only the effect for clothing and footwear can be considered significant,  also by considering the pre-trend slope as discussed in the Supporting Information.  Thus,  concerning $T_{3}$,  we considered not null,  in the scenario simulations,  only the clothing and footwear sectors.  By inspecting the Tab.  \ref{selling_DID_table_delta} we see that the most severe damage of lock-down hit the clothing,  and the footwear.  On the other side the Household kids and the Household tools seems lesser affected by the lock-down.  The same DID regression was used  for the subsidies,  using as control group the extraordinary solidarity subsidies (cassa integrazione straordinaria solidarietà): the plot of the data and the DID coefficients are reported in Fig. \ref{CIG_sector_plots} and in Tab \ref{CIG_DID}.  Also in this case the placebo tests were successfully performed.  An interesting insight is also provided by the inspection of the extraordinary subsidies for renovation ( \textit{cassa integrazione ristrutturazione}): these subsidies,  as compared to the solidarity ones,  seems much more sensitive with respect to external shock in particular for trade sector.

@@

\begin{figure}
     \centering
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Figures/CIG_effect_1_1_1.pdf}
         \caption{Active PM} 
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Figures/CIG_1.5_1.5_1.5.pdf}}
         \caption{Lazy PM}
     \end{subfigure}
\caption{Comparison between the subsides effects for the scenarios where an active vs lazy PM is considered.  These effects were calculated by running a DID regression for the different national lock-down imposed during March-June 2020 and then multiplying the number of lock-down weeks, obtianed from the SIRD scenario ,with the coefficients obtained from the DID (divided by 4).  The error bars, although are visible because are too reduced,  were calculated by considering the error propagation}
\label{CIG_sector_effects}
\end{figure}

>>>>

Among the very large number of compartmental model available in the literature \cite{vynnycky2010introduction} we considered,  as the simulator of the epidemic diffusion the simplest one: the Susceptible-Infectious-Recovered-Dead (SIRD).  Our choice is motivated by the fact that this relative simple model provides the gross features of an epidemic \cite{fernandez2020estimating,al2020forecasting} with a relative small number of parameters\footnote{One in principle can consider a SIRD model,in which the parameters that are time-dependent, as done by Ferrari et al. in Ref. \cite{ferrari2020modelling} for the description of Italian situation.  On the other side it is possible to increase the complexity of the model with other compartments as done in the following paper \cite{giordano2020modelling} by Giordano et al.  Note that in this last case the resolution of 9 differential equation is required (accompanied by the estimation of a large number of parameters)}). The SIRD model, first proposed by Kermack and McKendrick in 1927 \cite{kermack1927contribution},  is given by the following set of differential equations\cite{vynnycky2010introduction}:  

@@

where S is the number of people that are still susceptible, I the number of people that are infected and R people that are recovered while D are people that are death.  N denotes the total population, that for the timing of this paper,  it will be considered fixed\footnote{Othervise, if longer horizontal timing is considered, it is necessary to consider also a source term for the births and a well term for the natural deaths. For further details see \cite{vynnycky2010introduction}}. On the other side $\beta$,$\gamma$ and $\mu$ are the parameters that shape the probability by which one individual in the model moves from a compartment to another: in particular $\beta$ is the probability to be infected, $\gamma $ the probability to recover and $\mu$ the probability to die (basically the lethality defined  as the probability to die conditionally to be ill).  Usually epidemiologist are interested in the ratio:  

@@

For each sales category we run the regression reported in the Eq.  \ref{DID-eq} on de-seasoned data reported in Fig \ref{Selling_plots} and \ref{Selling_plots_II}.  The coefficients obtained are reported in Tab.  \ref{selling_DID_table_delta} (the estimation of the other parameters is given in Supporting Information D).  As one can point out from the plots,  it is true that there is a pre-trend in the data before the event,  however as proved numerically in the Supporting Info B,  the slope of this pre-trend is up to two order of magnitude lesser with respect to the slope in the lock-down T$_{1}$ and $T_{2}$.  As consequence this pre-trend,  compared to the lock-down effect,  can be considered negligible.   In order to make a further control,  a placebo test was performed by choosing timing before the pandemic of COVID19.  As illustrated in the Supporting Info C,  this test was successful.  On the other side,  concerning $T_{3}$  (low lock-down),  only the effect for clothing and footwear can be considered significant,  also by considering the pre-trend slope as discussed in the Supporting Information.  Thus,  concerning $T_{3}$,  we considered not null,  in the scenario simulations,  only the clothing and footwear sectors.  By inspecting the Tab.  \ref{selling_DID_table_delta} we see that the most severe damage of lock-down hit the clothing,  and the footwear.  On the other side the Household kids and the Household tools seems lesser affected by the lock-down.  The same DID regression was used  for the subsidies,  using as control group the extraordinary solidarity subsidies (cassa integrazione straordinaria solidarietà): the plot of the data and the DID coefficients are reported in Fig. \ref{CIG_sector_plots} and in Tab \ref{CIG_DID}.  Also in this case the placebo tests were successfully performed.  An interesting insight is also provided by the inspection of the extraordinary subsidies for renovation ( \textit{cassa integrazione ristrutturazione}): these subsidies,  as compared to the solidarity ones,  seems much more sensitive with respect to external shock in particular for trade sector.

@@

\begin{figure}
     \centering
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Figures/CIG_effect_1_1_1.pdf}
         \caption{Active PM} 
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \includegraphics[width=\textwidth]{Figures/CIG_1.5_1.5_1.5.pdf}}
         \caption{Lazy PM}
     \end{subfigure}
\caption{Comparison between the subsides effects for the scenarios where an active vs lazy PM is considered. These effects were calculated by running a DID regression for the different national lock-down imposed during March-June 2020 and then multiplying the number of lock-down weeks, obtianed from the SIRD scenario ,with the coefficients obtained from the DID (scaled in order to obtain the week value). The error bars, although are visible because are too reduced,  were calculated by considering the error propagation}
\label{CIG_sector_effects}
\end{figure}
